---
title: "R Code & Output_RRWM Activity_Genan"
author: "Genan Hamad"
date: "2025-10-02"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Project title: RRWM Data Activity CAnD3
## Owner: Genan Hamad 
## Date: Oct 2, 2025
Citation of Data: Statistics Canada. 2019. Census of Population, 2016 [Canada] Public Use Microdata File (PUMF): Individuals File. Statistics Canada [producer and distributor], accessed September 10, 2021. ID: pumf-98M0001-E-2016-individuals.
The following tasks are included in this script:  
Create a subset of data from the Education and Income variables
Recode the Education variable 
Generate a descriptive statistics table for the new subset of data
Run a simple linear regression
```{r}
# Import and Read CSV data file using base R
housing_data <- read.csv("housing_data_pumf-98M0001-E-2016-individuals_F1.csv")
```


```{r}
# Preview data
head(housing_data)
```
## Variable selection

```{r}
# Create a new dataset with only Wages (income variable) and HDGREE (education variable) columns
data2 <- housing_data[c("Wages", "HDGREE")]

# View first few rows of data2
head(data2)
```
## recoding the HDGREE (education variable)

```{r}
# recode the HDGREE (education variable), where as the values (1, 2, 3, 4, 5, 6, 7) 
# are coded as "low_edu" and the values (8, 9, 10, 11, 12, 13) are coded as "high_edu"

library(dplyr)

data2 <- data2 %>%
  mutate(HDGREE = case_when(
    HDGREE %in% 1:7 ~ "low_edu",
    HDGREE %in% 8:13 ~ "high_edu",
    TRUE ~ NA_character_  # for unexpected/missing values
  ))

# show summary of data2
summary(data2)

# Preview data2
head(data2)
```

## Descriptive Statistics Table
```{r}
# generate a descriptive statistics table for data2   
# data2 contains:
# "Wages": numeric variable and "HDGREE": categorical variable ("low_edu" and "high_edu")

# Install required packages
install.packages("summarytools", repos = "https://cloud.r-project.org")
install.packages("officer", repos = "https://cloud.r-project.org")

# Load them
library(summarytools)
library(officer)

# Create a discriptive statistics table for data2, using dfSummary() 
desc_table <- dfSummary(data2)
dfSummary(data2)
```
## Regression Analysis

```{r}
# perform a simple linear regression using data2, where:
# Outcome variable: Wages (numeric)
# Predictor: HDGREE (categorical: "low_edu", "high_edu")

# Use HDGREE as a factor
data2$HDGREE <- factor(data2$HDGREE, levels = c("low_edu", "high_edu"))

# Run a simple linear regression analysis 
model <- lm(Wages ~ HDGREE, data = data2)

# View the regression model summary
summary(model)
